Genetic draft and valley crossing

### 2018-06-05 ms/valley_crossing.tex
edited through simulation methods

\newcommand{\comment}[2][noinline]{\todo[color=red,#1]{#2}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Living systems are characterized by complex adaptations, at least some of which have arisen by evolutionary paths whose intermediate states are neutral or even deleterious.
Such adaptations have been termed ``irreducibly complex", and the process by which they evolve is known as ``crossing a fitness valley".
Previous efforts have rigorously characterized the rate at which such complex adaptations evolve in populations of roughly equally fit individuals.
However, populations that are very large or have broad fitness distributions, such as many microbial populations, adapt quickly, which substantially alters the fate of individual mutations.
We investigate the rate at which irreducible complexity evolves in these rapidly adapting populations, focusing on asexual populations.
We confirm that rapid adaptation overall increases the time required to cross a valley but can make it easier for deeper valleys to be crossed.
\end{abstract}

@@

Valley crossing in asexual populations is increasingly well understood \citep{weissman_2009}.
If the number of mutants produced every generation is high, which occurs in large populations with high mutation rates, a complex adaptation requiring multiple mutations that is destined to fix will eventually be generated de novo. This is sometimes called deterministic fixation of the double mutant \citep{weissman_2009}.
In small populations, neutral or deleterious intermediate mutants generated by wild type individuals can drift to fixation. Then, the further beneficial mutations necessary for the adaptation can fix due to positive selection. This is the sequential fixation regime.
A third possibility occurs for intermediate population sizes where wild type individuals generate transient mutant subpopulations or ``bubbles".
These bubbles are ordinarily doomed to extinction due to drift, but additional mutations in a lucky bubble may generate a complex adaptation that sweeps to fixation; this process has been referred to as ``tunneling" \citep{iwasa_2004, weissman_2009}.
On its own, valley crossing becomes more likely as the population size increases \citep{weissman_2009}. However, recent work suggests that valley crossing relative to the fixation of a simple beneficial mutation is least likely at intermediate population sizes where tunneling occurs \citet{ochs_2015}. This suggests that tunneling may be the most difficult mode of valley crossing. 
% Valley crossing in sexually reproducing populations was by \citet{weissman_2010}, which found that the criterion $r < s$, with $r$ the recombination rate and $s$ the selective advantage of the complex adaptation, is the necessary criterion for recombination to be helpful in crossing a valley.
% Fairly low recombination rates can bring mutant subpopulations together, increasing the rate at which complex adaptations arise and sweep.
% High recombination rates, on the other hand, lower the valley crossing rate by forcing fit individuals to outcross with the less fit wild type.

@@

There are three components to the process of stochastic tunneling.
First, a single mutant lineage must appear, which occurs with rate $N\mu$.
If this lineage arises at time $t_0$ and persists until time $T$, then it gives rise to $W = int_{t_0}^T n(t) dt$ total individuals before going extinct, where $n(t)$ is the number of mutant individuals extant at time $t$.
We refer to such a short-lived mutant lineage, and all the individuals generated within that lineage, as a mutant ``bubble": the time integrated number of individuals $W$ is the ``weight" of the bubble.
During the lifetime of a bubble, the complex adaptation must appear and establish; that is, it must rise to a high enough frequency that fixation is almost guaranteed, which occurs at rate $\mu \pfix$ where $\pfix$ is the probability that the double mutant reaches fixation from a single initial individual.
If it establishes, the complex adaptation will quickly sweep to fixation.
The first two steps can be folded in together and considered as one process, so that what is relevant is the expected time until a single mutant lineage arises that is destined to produce a successful double mutant.
In this way, $\mathbb{E}\left[ T \right] = \mathbb{E} \left[ T_0 + T_1 \right]$, where $T_0$ is the time to the first successful bubble--the first that produces a double mutant that is destined for fixation--and $T_1$ is the time required for the double mutant to sweep.
We assume that $T_0 \gg T_1$, so that the wait time is dominated by the first term
\comment{JVC: justify this briefly}.

@@

\comment[inline]{
  \begin{itemize}      
  \item  This section is still too technical. For example, the reader here doesn't need to know what  ``weak driver'' is, or if they do, it should be more fully explained and used more than just in what is effectively a background section. 
  \item I think having the PDE for the fixation probability is ok, but you don't need the details about the limits necessary for deriving it. The point of keeping it more informational for the reader. 
  \item Also, avoid introducing new variables, like $D$, and keep the existing notation, like $\pfix$, instead of changing to $p$ since the simplicity doesn't save much time (we only use the equation once) and the extra notation could be confusing. 
  \item Part of the motivation for introducing the PDE in the first place is to then say what the results of analyzing it, namely what the fixation probability is under draft with fitness background $x$. You explain this a little bit, but you can more fully describe how the $\pfix$ increases exponentially then linearly. 
        \item Also, I don't think you mention how the valley depth $\delta$ and peak height $s$ related to the background fitness $x$ and fitness variance $\sigma$. This should probably come up here, but also probably in the first paragraph of the ``Mathematical background''
        \end{itemize}
      }
{\color{red}
We begin by considering the dynamics of $\pfix(x,s)$ before moving on to consider the distribution of bubble sizes.
We zero in on the ``weak driver" model, in which background mutations are frequent but of weak effect compared to $\sigma$.
In this case, the background beneficial mutation rate $U_b$ and mean fitness effect $\bar{\epsilon}$ turn out not to matter: what is relevant is the product $D = U_b \bar{\epsilon}^2/2$.
Taking the limit $U_b \to \infty$ and $\bar{\epsilon} \to 0$ yields the ``infinitesimal" model.
\citet{hallatschek_2011} and \citet{good_2014} showed that the fixation probability in this regime obeys the differential equation 
\begin{equation}
v\partial_x p(x,s) = x p(x,s) + D\partial_x^2 p(x,s) - \frac{p(x,s)^2}{2},
\end{equation}
in which we have assumed that $v \approx \sigma^2$, which is tantamount to claiming that mutational effects are less important than selection in determining the advance of the mean fitness.
The fixation probability then increases linearly in $x$ above some critical value $x_c$ and remains small otherwise, with $x_c \approx \sigma^4/4D$: when $\sigma$ is large, only mutations in the nose stand any chance of fixing unless they can be dragged to higher fitness by a background mutations (large $D$).
}

@@

To analyze our model, we perform forward simulations using a customized version of \texttt{FFPopSim} \citep{zanini_2012}, a discrete time forward simulation package that implements a modified Wright-Fisher model.
The simulation code is available via \texttt{GitHub}.
In \texttt{FFPopSim}, individuals are organized by ``clones" (sets of individuals with the same genome): each individual gives rise to a Poisson distributed number of offspring, with the mean dependent on the relative fitness.
The mean offspring number is further adjusted to keep the population roughly at a pre-specified carrying capacity.
We initialize a wild type population of $N$ individuals consisting of haploid genomes of length $L = 200$.
Genetic loci are partitioned into two groups, the two ``focal" loci (at which the epistatic alleles will segregate) and ``background" loci (which are responsible for the underlying fitness distribution).
For the background loci, we make use of a modified ``infinite sites" model: any time a locus becomes monomorphic, a mutation at that locus is injected into a random individual in the population.
In this way, the population experiences a constant influx of beneficial mutations at the background loci.
These mutations' fitness effects are drawn from an exponential distribution; the background fitness variance is set to a constant value $\sigma^2$ by manually adjusting the selection coefficients every generation (they are multipled by the current variance and divided by $\sigma^2$).
In this way, the background mutation rate $U$ and average fitness effect $\bar{\epsilon}$ are not parameters of our simulation model but must be directly measured from simulations: they are constrained by $L$ and $\sigma$.
We estimate them by manually counting the number of injected mutations every generation and averaging over the fitness effect at each locus.

We set the ``focal" loci (the loci at which the epistatic alleles segregate) to be at positions $L/4$ and $3L/4$.
Simulations are allowed to proceed for an ``equilibration" time of $N/10$ generations to allow sufficient genetic diversity to be introduced.
During this time, the mutation rate $\mu$ for the focal loci is set to zero.
After equilibration, we set a per-site mutation rate $\mu$ for the focal loci (which controls both forward and backward mutations), with single mutant fitness disadvantage $\delta$ and double mutant advantage $s$, consistent with our analytical model.
When the double mutant reaches frequency $0.5$, we consider it to have fixed.
We then set the focal loci allele frequencies back to zero, randomize the selection coefficients for the genetic background by drawing anew from an exponential distribution, and allow the population to equilibrate for another $N/10$ generations.
This offers some assurance that independent valley crossing trials are indeed independent despite occurring in the same simulation run.

We partitioned our simulations as follows.
We first studied the effect of increasing $\sigma$ in what \citet{weissman_2009} describes as the sequential fixation, stochastic tunneling, semi-deterministic tunneling, and sequential fixation regimes.
In the case of sequential fixation, the valley crossing time is dominated by the time until a deleterious mutant that is destined to fix arises: once it does, it does so in order $T_{\mathrm{MRCA}}$ generations, which is of order $N$ for low $N\sigma$ and scales weakly with $N$ for higher $N\sigma$ \citep{neher_hallatschek_2013}.
In the drift limit, sequential fixation applies when $N \ll \mathrm{min}(1/\delta, 1/\sqrt{\mu s})$, so that a successful mutant is unlikely to arise until the intermediate has fixed, but the intermediate's fitness disadvantage is too weak for selection to inhibit it.
Next, we considered the stochastic tunneling regime.
Tunneling becomes relevant when $N$ is too large for sequential fixation, i.e., the tunneling probability becomes more significant than the probability that the intermediate will fix: this mandates $N \gtrsim 1/\sqrt{\mu s}$.
It also requires that $\ll 1/\mu$, meaning that only one mutant lineage is likely to be extant at any given time and, if they do co-occur, are unlikely to interfere with each other or substantially affect the mean fitness.
This happens to be the necessary condition for a birth-death process to be an appropriate model.
By increasing $N$ far above the $1/\mu$ limit, multiple deleterious mutants are likely to occur every generation, reaching mutation-selection balance at low frequency and providing many opportunities for a lucky double mutant to appear (so-called ``semi-deterministic tunneling").
At still higher population sizes ($\gtrsim s/\mu^2$), multiple double mutants are likely to appear every generation, and their dynamics can be modeled as deterministic.

Finally, we compared the dynamics of valley crossing versus sweeping mutations.
We performed simulations with a sweeping focal beneficial mutation, with selection coefficient $s_{\mathrm{sweep}}$, at position $L/2$, with the background loci kept polymorphic as in the valley crossing simulations.
We recorded the time for a beneficial mutation to arise and reach frequency $0.5$, considering it to have fixed at that point.
We then computed the ratio of the sweep time to the valley crossing time for a comparable complex adaptation, with valley depth $\delta$ and larger fitness advantage $s_{\mathrm{valley}}$.
To determine the effect of $\sigma$ on this ratio, we performed simulations at both high and low $\sigma$ and computed the ratio of these ratios: the result is a measure of the extent to which rapid adaptation favors valley crossing over simple sweeps.

@@

%%% from into but more relevant here
% In sexual populations, the accuracy of this description depends on the recombination model.
% When recombination is ``free", meaning that child genomes are assembled by randomly shuffling those of their parents, there is a chance for new alleles on weak fitness backgrounds to jump onto fitter ones near the nose of the distribution \citep{neher_shraiman_2011}.
% Competition between recombination, which breaks up linkage, and adaptation, which creates it, determines whether draft or drift is more important.
% When recombination occurs via crossovers, genomes can be partitioned into effectively asexual ``blocks", whose size depends on the per-site crossover rate and the total fitness variation segregating in the population.
% If each block harbors some portion $\sigma_b$ of the total fitness standard deviation $\sigma$, then the conditions $N\sigma_b \gg 1$ and $N\sigma_b \ll 1$ correspond to draft and drift dominance, respectively \citep{neher_kessinger_2013}.

@@

%In the case where $\rho = 0$ and the fitness variance $\sigma^2$ is small ($N\sigma \ll 1$), the rate of valley crossing agrees with \citet{weissman_2009}: when $\rho > 0$, the rate agrees with \citet{weissman_2010}.
%This is because, provided that $N\sigma \ll 1$, genetic drift is the major evolutionary process affecting the frequency of deleterious intermediates, and the fixation probability of the complex adaptation simply scales with $s$.
%We turn to the regime where $N\sigma \gg 1$, meaning that draft is more important than drift, but the fixation probability of the full complex adaptation is suppressed.
%By the central limit theorem, the background loci create a background fitness wave whose shape is roughly Gaussian.
%We assume that the constant influx of weak effect beneficial mutations keeps the fitness variance $\sigma^2$ roughly constant.
%We sketch some heuristics in this limit before providing a more formal analysis.

% \iffalse
% First, we consider the case where $\rho = 0$ and focus on what \citet{weissman_2009} refer to as the process of ``stochastic tunneling", where $N\mu \ll 1$.
% As we shall see, the results in the $\rho = 0$ and low $\rho$ limits can easily be extended to sexually recombining populations: moreover, understanding the rate of tunneling will allow us to easily examine the rate at which sequential fixation and deterministic fixation occur.
% In stochastic tunneling, a deleterious intermediate arises (at rate $N\mu$), and its lineage persists over a short time $T$, forming a transient ``bubble".
% If the number of individuals in the lineage is given by $n(t)$, then the total number of individuals in the ``bubble" is $w = \int_{t_0}^T n(t) dt$.
% We refer to $w$ as the ``weight" of the bubble.
% The overall rate of valley crossing becomes $1/N\mu p$, with $p$ the probability that a particular bubble will give rise to a successful double mutant.
% Thus, given that a deleterious intermediate appears, it has $w$ ``chances" to generate a successful complex adaptation, which it does at rate $\sim \mu s$: double mutants arise at rate $\mu$ and have a probability $\sim s$ of fixing.
% The probability of achieving a weight of at least $w$ scales as $1/\sqrt{w}$ (or, equivalently, $P(w) \sim w^{-3/2}$), meaning that with probability $\sqrt{\mu s}$, a bubble gives rise to a successful single mutant.
% This probability drops off rapidly at $1/\delta^2$, so for higher values of $\delta$ (so called ``deleterious tunneling"), the rate at which new complex adaptations are generated starts to fall: the size of the mutant subpopulation is bounded from above by roughly $1/\delta$.

% In the high $N\sigma$ regime, the argument is much the same, but the underlying dynamics are different.
% In stark contrast to the drift case, the fate of a complex adaptation depends strongly on the fitness $x$ of the genetic background on which it appears.
% Individuals with fitness lower than some critical fitness $x_c$ are almost guaranteed to go extinct.
% This means the fixation probability of the complex adaptation no longer scales with $s$, and $P(w) \sim w^{-2}$ instead of $w^{-3/2}$, meaning that the probability of reaching weight $w$ is $\sim 1/w$.
% However, individuals with higher fitnesses can be destined for fixation, even if they are carrying a significant load of deleterious mutations.
% This means the dependence on $\delta$ is significantly lessened: even deep fitness valleys can be crossed by a rapid enough wave.
% (Still needed: expression for fixation probability of beneficial mutant--this should be easy enough to obtain.
% Can start with equation 3 of \citet{neher_shraiman_2011} and drop the recombination term, which is what I'm doing now.
% This is tantamount to a ``mean field" approximation but might still work out okay.
% Other possible approach is to use something like the expression of  \citet{desai_fisher_2007} for the fixation probability and rescale it to a traveling wave model.
% Once we have the fixation probability, studying sequential and deterministic fixation will be easier than studying tunneling, since sequential fixation is essentially two different fixation events.)

% One difficulty in computing the rate of tunneling in the high $N\sigma$ case is that the expected bubble size and fixation probability will both turn out to depend on the background fitness of an individual.
% That is to say, the total rate of valley crossing will be given by $\int P(x) \pfix(x,s) \mathbb{E}\left[ w(x,\delta) \right] dx$, with $P(x)$ the (presumably Gaussian) background fitness distribution.
% One cannot simply compute $\pfix$ and the weight distribution by themselves but rather must convolute both of them over the background fitness.
% This is intuitive: individuals near the nose of the distribution will tend to give rise to longer-lived lineages, and because they are near the nose, the complex adaptation will therefore likewise be more likely to fix.
% We proceed to derive both $\pfix(x,s)$ and $P(w)$ in the zero recombination limit.

% The establishment probability $p(x,s)$ obeys the PDE
% \begin{equation}
% v \partial_x p(x,s) = (x+s)p(x,s) - (1+x+s)p(x,s)^2.
% \end{equation}
% To see this, consider the generic expression for the fixation probability
% \begin{equation}
% p(X,t-dt) = p(X,t)-dt(D+B(X,t))p(X,t) + dt(B(X,t)(2p(X,t)-p(X,t)^2)
% \end{equation}
% \citep{barton_1995}. Taking the limit as $dt \to 0$ and inserting $D = 1, B = 1 + X -\bar{X}(t)+s$ yields a PDE for $p(X,t)$.
% Defining $x = X - vt$, we then have $\partial_t p(X-vt) = -v\partial_x p(x)$.

% This is a Bernoulli equation and is readily solved by substituting $u(x,s) = 1/p(x,s)$. We then have that $\partial_x u(x,s) = -\frac{\partial_x p(x,s)}{p(x,s)^2}$, so that
% \begin{equation}
% \partial_x u(x,s) = -\frac{(x+s)}{v}u(x,s) - \frac{1+x+s}{v}.
% \end{equation}
% The use of the integrating factor $e^{\int \frac{x+s}{v} dx} = Ce^{\frac{(x+s)^2}{2v}}$ yields
% \begin{equation}
% Ce^{\frac{(x+s)^2}{2v}} (\partial_x u + \frac{x+s}{v}u) = -Ce^{\frac{(x+s)^2}{2v}}\frac{1+x+s}{v}.
% \end{equation}
% The left side is simply $\partial_x(Ce^{\frac{(x+s)^2}{2v}}u)$, so (after dividing $C$ out) we can integrate both sides, then multiply by $e^{\frac{-(x+s)^2}{2v}}$ to obtain
% \begin{equation}
% u(x,s) =  -e^{\frac{-(x+s)^2}{2v}} \frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x+s}{\sqrt{2v}}}{\sqrt{v}}) - 1 + C(e^{\frac{-(x+s)^2}{2v}}),
% \end{equation}
% with $C$ a constant. Of course, $p = 1/u$.

% Unfortunately it is not clear how to get rid of $C$, as there is no obvious ``initial condition".
% The most we are likely to know about $p$ is that $p$ approaches $0$ in the limit of low $x$, but behaves $\sim s$ in the limit of high $x$ before saturating at $1$ (however, there are unlikely to be any individuals in the population at such high $x$).
% One possibility is to impose an \emph{ersatz} ``solvability condition": surely the fixation probability of a neutral mutation averaged over the entire population must equal $1/N$. We would then have
% \begin{equation}
% \int_{-\infty}^{\infty} P(x)p(x,0)dx = \frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty} \frac{e^{-x^2/2\sigma^2}}{-e^{-x^2/2\sigma^2} \frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x}{\sqrt{2\sigma^2}})}{\sigma} - 1 + Ce^{-x^2/2\sigma^2}}dx = \frac{1}{N}.
% \end{equation}
% Multiplying the top and bottom of the integrand by $e^{x^2/2\sigma^2}$ would clean this up, yielding
% \begin{equation}
% \frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty} \frac{1}{-\frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x}{\sqrt{2\sigma^2}})}{\sigma} - e^{x^2/2\sigma^2} + C}dx = \frac{1}{N}.
% \end{equation}
% Unfortunately this integral has an imaginary error function in the denominator, so there is not likely to be a closed form expression: we could explore it numerically and see where it peaks, however.

% Another idea would be to consider an alternative way of solving the differential equation for $p$.
% Return to
% \begin{equation}
% v \partial_x p(x,s) = (x+s)p(x,s) - (1+x+s)p(x,s)^2.
% \end{equation}
% Suppose we make the \emph{ansatz} that $x+s \ll 1$, and define $\chi = x/\sigma$, $\tilde{s} = s/\sigma$, and $\tilde{p}(\chi,\tilde{s}) = p(x,s)/\sigma$. Then
% \begin{equation}
% \partial_\chi \tilde{p}(\chi,\tilde{s}) = (\chi + \tilde{s})\tilde{p}(\chi,\tilde{s}) - \tilde{p}(\chi,\tilde{s})^2.
% \end{equation}
% Now define $q(\chi,\tilde{s})$ such that $\tilde{w}(\chi,\tilde{s}) = \partial_x \log q(\chi,\tilde{s})$.
% Then
% \begin{equation}
% \partial_\chi^2 q(\chi,\tilde{s}) - (\chi+\tilde{s})\partial_\chi q(\chi, \tilde{s}) = 0.
% \end{equation}
% Defining $\vartheta = \chi + \tilde{s}$ and $q(\vartheta) = e^{\vartheta^2/4}h(\vartheta)$ yields
% \begin{equation}
% \partial_\vartheta^2 h(\vartheta) - (\frac{1}{2} + \frac{\theta^2}{4})h(\vartheta) = 0,
% \end{equation}
% which is the parabolic cylinder equation.
% We then need to find the form of $q$ that has the correct asymptotics.
% This does have a solution:
% \begin{equation}
% q(\chi) = \int_0^\infty d\lambda e^{\vartheta \lambda - \lambda^2/2}/\lambda.
% \end{equation}

>>>>

\newcommand{\jcomment}[2][noinline]{\todo[size=\tiny,author=JVC,color=red,#1,caption={},fancyline]{#2}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Living systems are characterized by complex adaptations where multiple mutations are required for the function of the adaptation. Empirical studies of the fitness landscape that results from the many possible mutations in a gene region reveal multiple possible fitness peaks and valleys that connect them. Thus, it is possible that some complex adaptations have arisen by evolutionary paths whose intermediate states are neutral or even deleterious.
Traversing these evolutionary paths when intermediates are deleterious is known as ``crossing a fitness valley".
Previous efforts at studying this problem have rigorously characterized the rate at which such complex adaptations evolve in populations of roughly equally fit individuals.
However, populations that are very large or have broad fitness distributions, such as many microbial populations, adapt quickly, which substantially alters the fate of individual mutations due to the action of genetic draft.
We investigate the rate at which complex adaptations evolve in these rapidly adapting populations in regions without recombination.
We confirm that rapid adaptation overall increases the time required to cross a valley but can make it easier for deeper valleys to be crossed relative to the time required for single beneficial mutations to sweep to fixation.
\end{abstract}

@@

Valley crossing in asexual populations is increasingly well understood \citep{weissman_2009}.
If the number of mutants produced every generation is high, which occurs in large populations with high mutation rates, a complex adaptation requiring multiple mutations that is destined to fix will eventually be generated de novo. This is sometimes called deterministic fixation of the double mutant \citep{weissman_2009}.
In small populations, neutral or deleterious intermediate mutants generated by wild type individuals can drift to fixation. Then, the further beneficial mutations necessary for the adaptation can fix due to positive selection. This is the sequential fixation regime \citep{weissman_2009}.
A third possibility occurs for intermediate population sizes where wild type individuals generate transient mutant subpopulations or ``bubbles".
These bubbles are ordinarily doomed to extinction due to drift, but additional mutations in a lucky bubble may generate a complex adaptation that sweeps to fixation; this process has been referred to as ``tunneling" \citep{iwasa_2004, weissman_2009}.
On its own, valley crossing becomes more likely as the population size increases \citep{weissman_2009}. However, recent work suggests that valley crossing relative to the fixation of a simple beneficial mutation is least likely at intermediate population sizes where tunneling occurs \citet{ochs_2015}. This suggests that tunneling may be the most difficult mode of valley crossing. 
% Valley crossing in sexually reproducing populations was by \citet{weissman_2010}, which found that the criterion $r < s$, with $r$ the recombination rate and $s$ the selective advantage of the complex adaptation, is the necessary criterion for recombination to be helpful in crossing a valley.
% Fairly low recombination rates can bring mutant subpopulations together, increasing the rate at which complex adaptations arise and sweep.
% High recombination rates, on the other hand, lower the valley crossing rate by forcing fit individuals to outcross with the less fit wild type.

@@

There are three components to the process of stochastic tunneling.
First, a single mutant lineage must appear, which occurs with rate $N\mu$.
If this lineage arises at time $t_0$ and persists until time $T$, then it gives rise to $W = int_{t_0}^T n(t) dt$ total individuals before going extinct, where $n(t)$ is the number of mutant individuals extant at time $t$.
We refer to such a short-lived mutant lineage, and all the individuals generated within that lineage, as a mutant ``bubble": the time integrated number of individuals $W$ is the ``weight" of the bubble.
During the lifetime of a bubble, the complex adaptation must appear and establish; that is, it must rise to a high enough frequency that fixation is almost guaranteed, which occurs at rate $\mu \pfix$ where $\pfix$ is the probability that the double mutant reaches fixation from a single initial individual.
If it establishes, the complex adaptation will quickly sweep to fixation.
The first two steps can be folded in together and considered as one process, so that what is relevant is the expected time until a single mutant lineage arises that is destined to produce a successful double mutant.
In this way, $\mathbb{E}\left[ T \right] = \mathbb{E} \left[ T_0 + T_1 \right]$, where $T_0$ is the time to the first successful bubble--the first that produces a double mutant that is destined for fixation--and $T_1$ is the time required for the double mutant to sweep.
We assume that $T_0 \gg T_1$, so that the wait time is dominated by the first term
\jcomment{justify this briefly}.

@@

\jcomment[inline]{
  \begin{itemize}      
  \item  This section is still too technical. For example, the reader here doesn't need to know what  ``weak driver'' is, or if they do, it should be more fully explained and used more than just in what is effectively a background section.
  \item I think having the PDE for the fixation probability is ok, but you don't need the details about the limits necessary for deriving it. The point of keeping it more informational for the reader. 
  \item Also, avoid introducing new variables, like $D$, and keep the existing notation, like $\pfix$, instead of changing to $p$ since the simplicity doesn't save much time (we only use the equation once) and the extra notation could be confusing. 
  \item Part of the motivation for introducing the PDE in the first place is to then say what the results of analyzing it, namely what the fixation probability is under draft with fitness background $x$. You explain this a little bit, but you can more fully describe how the $\pfix$ increases exponentially then linearly. 
  \item Also, I don't think you mention how the valley depth $\delta$ and peak height $s$ related to the background fitness $x$ and fitness variance $\sigma$. This should probably come up here, but also probably in the first paragraph of the ``Mathematical background''
  \end{itemize}
}
{\color{red}
We begin by considering the dynamics of $\pfix(x,s)$ before moving on to consider the distribution of bubble sizes.
We zero in on the ``weak driver" model, in which background mutations are frequent but of weak effect compared to $\sigma$.
In this case, the background beneficial mutation rate $U_b$ and mean fitness effect $\bar{\epsilon}$ turn out not to matter: what is relevant is the product $D = U_b \bar{\epsilon}^2/2$.
Taking the limit $U_b \to \infty$ and $\bar{\epsilon} \to 0$ yields the ``infinitesimal" model.
\citet{hallatschek_2011} and \citet{good_2014} showed that the fixation probability in this regime obeys the differential equation 
\begin{equation}
v\partial_x p(x,s) = x p(x,s) + D\partial_x^2 p(x,s) - \frac{p(x,s)^2}{2},
\end{equation}
in which we have assumed that $v \approx \sigma^2$, which is tantamount to claiming that mutational effects are less important than selection in determining the advance of the mean fitness.
The fixation probability then increases linearly in $x$ above some critical value $x_c$ and remains small otherwise, with $x_c \approx \sigma^4/4D$: when $\sigma$ is large, only mutations in the nose stand any chance of fixing unless they can be dragged to higher fitness by a background mutations (large $D$).
}

@@

To analyze our model, we perform forward simulations using a customized version of \texttt{FFPopSim} \citep{zanini_2012}, a discrete time forward simulation package that implements a modified Wright-Fisher model.
The simulation code is available via \texttt{GitHub}.
In \texttt{FFPopSim}, individuals are organized by ``clones" (sets of individuals with the same genome): each individual gives rise to a Poisson distributed number of offspring, with the mean dependent on the relative fitness.
The mean offspring number is further adjusted to keep the population roughly at a pre-specified carrying capacity.
We initialize a wild type population of $N$ individuals consisting of haploid genomes of with $L = 200$ loci.
Genetic loci are partitioned into two groups, the two ``focal" loci where the epistatic alleles $A$ and $BB$ will segregate and the remaining ``background" loci that are responsible for the underlying fitness distribution.
We set the focal loci to be at positions $L/4$ and $3L/4$.
For the background loci, we make use of a modified infinite sites model \citep{Kimura:1971,Watterson:1975}: any time a locus becomes monomorphic, a mutation at that locus is injected into a random individual in the population.
In this way, the population experiences a constant influx of beneficial mutations at the background loci.
These mutations' fitness effects are drawn from an exponential distribution; the background fitness variance is set to a constant value $\sigma^2$ by manually adjusting the selection coefficients every generation (they are multipled by the current variance and divided by $\sigma^2$).
In this way, the mutation rate $U$ \jcomment{Is this $U_{B}$ or $U$?} and average fitness effect $\bar{\epsilon}$ at the background loci \jcomment{this should be defined clearly either above in this section or in the math background section (it appears in that section but is not explicitly defined)} are not parameters of our simulation model but must be directly measured from simulations; i.e., they are constrained by $L$ and $\sigma$.
We estimate them by manually counting the number of injected mutations every generation and averaging over the fitness effect at each locus.

Simulations are allowed to proceed for an equilibration time of $N/10$ generations to allow sufficient genetic diversity to be introduced.
During this time, the mutation rate $\mu$ for the focal loci is set to zero.
After equilibration, we set a per-site mutation rate $\mu$ for the focal loci (allowing forward and backward mutations) with single mutant fitness disadvantage $\delta$ and double mutant advantage $s$.
When the double mutant reaches frequency $0.5$, we consider it to have fixed.
We then set the focal loci allele frequencies back to zero, randomize the selection coefficients for the genetic background by drawing anew from an exponential distribution, and allow the population to equilibrate for another $N/10$ generations.
This ensures that independent valley crossing trials are independent despite occurring in the same simulation run.

We partitioned our simulations as follows.
We first studied the effect of increasing $\sigma$ in the sequential fixation, stochastic tunneling, semi-deterministic tunneling, and sequential fixation regimes \citep{weissman_2009}.
\jcomment{commented out some text here; put back into the results if the pattern described is evident there}
% In the case of sequential fixation, the valley crossing time is dominated by the time until a deleterious mutant that is destined to fix arises: once it does, it does so in order $T_{\mathrm{MRCA}}$ generations, which is of order $N$ for low $N\sigma$ and scales weakly with $N$ for higher $N\sigma$ \citep{neher_hallatschek_2013}.
For small populations, sequential fixation applies when $N \ll \min(1/\delta, 1/\sqrt{\mu s})$, so that a successful mutant is unlikely to arise until the intermediate has fixed, but the intermediate's fitness disadvantage is too weak for selection to inhibit it.
Tunneling becomes relevant when $N$ is too large for sequential fixation, i.e., the tunneling probability becomes more significant than the probability that the intermediate will fix: this mandates $N \gtrsim 1/\sqrt{\mu s}$.
It also requires that $\ll 1/\mu$ \jcomment{$N$ missing here?}, meaning that only one mutant lineage is likely to be extant at any given time and, if lineages do co-occur, they are unlikely to interfere with each other or substantially affect the mean fitness.
This happens to be the necessary condition for a birth-death process to be an appropriate model.\jcomment{is this previous sentence necessary?}
By increasing $N$ far above the $1/\mu$ limit, multiple deleterious mutants are likely to occur every generation, reaching mutation-selection balance at low frequency and providing many opportunities for a lucky double mutant to appear; this is semi-deterministic tunneling.
At still higher population sizes ($N \gtrsim s/\mu^2$), multiple double mutants are likely to appear every generation, and their dynamics can be modeled as deterministic.

Finally, we compared the dynamics of valley crossing versus sweeping mutations.
We performed simulations with a sweeping focal beneficial mutation, with selection coefficient $s_{\mathrm{sweep}}$, at position $L/2$, with the background loci kept polymorphic as in the valley crossing simulations.
We recorded the time for a beneficial mutation to arise and reach frequency $0.5$, considering it to have fixed at that point.
We then computed the ratio of the sweep time to the valley crossing time for a complex adaptation, with valley depth $\delta$ and larger fitness advantage $s_{\mathrm{valley}}$.
To determine the effect of $\sigma$ on this ratio, we performed simulations at both high and low $\sigma$ and computed the ratio of these ratios: the result is a measure of the extent to which rapid adaptation increases the likelihood of valley crossing relative to the speed of simple sweeps.

@@

\jcomment{check out some text that was from the intro that I've commented out and see if it fits here}
%%% from into but more relevant here
% In sexual populations, the accuracy of this description depends on the recombination model.
% When recombination is ``free", meaning that child genomes are assembled by randomly shuffling those of their parents, there is a chance for new alleles on weak fitness backgrounds to jump onto fitter ones near the nose of the distribution \citep{neher_shraiman_2011}.
% Competition between recombination, which breaks up linkage, and adaptation, which creates it, determines whether draft or drift is more important.
% When recombination occurs via crossovers, genomes can be partitioned into effectively asexual ``blocks", whose size depends on the per-site crossover rate and the total fitness variation segregating in the population.
% If each block harbors some portion $\sigma_b$ of the total fitness standard deviation $\sigma$, then the conditions $N\sigma_b \gg 1$ and $N\sigma_b \ll 1$ correspond to draft and drift dominance, respectively \citep{neher_kessinger_2013}.

@@
